Excitonic Energy Transfer Couplings at Short Distances: Efficient Transfer-Integral Fragmentation Approach Based on Effective One-Electron Operators

### 2020-02-21 papers/4_eet.I.Theory/paper-r0.tex

 \mathscr{F} \approx \mathscr{F}^{A}_0 + \mathscr{F}^{B}_0 \;,

@@

@@

 \mathscr{F}^{X}_0 = \hat{T}^{X} + \hat{V}^{X}_{\rm nuc} + \hat{J}({\bf P}^{X}_g) - \frac{1}{2} \hat{K}({\bf P}^{X}_g) \;.

@@

Note that, without loss of generality, one can rewrite Eq.~\eqref{e:v0-et1}

@@

 \tbraket{\phi_L^A}{\mathscr{F}}{\phi_L^B} \\

@@

 \bra{\phi_L^A} \left\{ \mathscr{F}_0^A - \hat{v}^A_{HH} \right\}

@@

 \mathscr{F}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B;{\rm ET}}_{\eta;L} \ket{\eta} 

@@

Similar considerations apply to the hole transfer (HT) matrix elements.
The expression
%
\begin{multline}\label{e:v0-ht1}
 V^{{\rm HT1},(0)} \equiv \left< \Phi_1 \vert \mathscr{H} \vert \Phi_4 \right> \approx 
 t_{H\rightarrow L}^A \times \Big\{ 
-\big(\phi_H^A \big| \mathscr{F} \big| \phi_H^B \big) \\
   + 2 \big( \phi_H^A \phi_L^A \big| \phi_L^A \phi_H^B \big) - \big( \phi_H^A \phi_H^B \big| \phi_L^A \phi_L^A \big) 
 \Big\} 
\end{multline}
%
can be recast as
%
\begin{multline}\label{e:v0-ht1-oep}
 V^{{\rm HT1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
 \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;HL} +
 \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;H}
 \Big\} \;,
\end{multline}
%
with the effective potential matrices given by
%
%
\begin{subequations}
\begin{align}
 \bra{\phi_H^A} \left\{ -\mathscr{F}_0^A - \hat{v}^A_{LL} \right\}
 + \bra{\phi_L^A} 2 \hat{v}^A_{HL} 
 &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT}}_{\xi,HL} \\
 -\mathscr{F}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm HT}}_{\eta,H} \ket{\eta} 
\end{align}
\end{subequations}
%
The twin term HT2 is accordingly
%
\begin{multline}\label{e:v0-ht2-oep}
 V^{{\rm HT2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
 \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;HL} +
 \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;H}
 \Big\} \;.
\end{multline}
%
As can be seen, to compute Hamiltonian matrix elements
associated with the electron and hole transfer, eight in total
different OEP matrices need to be pre\hyp{}computed and stored in a file.
Note that each of these matrices is actually just a vector of length
equal to the size of the auxiliary basis set chosen. Therefore,
it can be predicted that the computational cost of evaluation of the OEP\hyp{}based Hamiltonian
matrix elements is negligible as compared to the original TI/CIS method.
In Appendix~\ref{a:gdf-formulae}, explicit working formulae
for the OEP matrix elements are given in terms of the AO's.

@@

As developed in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.XXXX}, 
the matrix elements of the OEP operators
are computed from the generalized density fitting scheme, according to which
%
\begin{equation}\label{e:a:voep}
 V^X_{\xi, M} = \sum_{\xi'\in X}^{\rm DF} 
                \sum_{\varepsilon\in X}^{\rm RI}
                [{\bf R}^{-1}]_{\xi\xi'} R_{\xi'\varepsilon} H^{X;M}_{\varepsilon; N} \;,
\end{equation}
%
where
%
\begin{equation}
 H^{X;M}_{\varepsilon; N} = \sum_{\varepsilon'\in X}^{\rm RI}
                        [{\bf S}^{-1}]_{\varepsilon\varepsilon'} a_{\varepsilon';N}^{X; M} \;.
\end{equation}
%
In the above equations, the auxiliary matrices are defined by
%
\begin{equation}
 S_{\eta\xi}  = \int 
                       \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) 
                 d{\bf r} 
\end{equation}
%
and
%
\begin{equation}
 R_{\eta\xi}  = \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
                 d{\bf r}_1 d{\bf r}_2 \;.
\end{equation}
%
The elements of vector ${\bf a}^{X;M}_N$ for $M=$ ET1, ET2, HT1 and HT2 and $N=$ H, L and HL are given by
%
\begin{subequations}\label{e:a:aoep}
\begin{align}
%
a_{\alpha;L}^{X;{\rm ET}} &= \sum_\beta C_{\beta L}^X F_{\alpha\beta}^X \;,\\
%
a_{\alpha;HL}^{X;{\rm ET}} &= a_{\alpha;L}^{X;{\rm ET}} \nonumber \\
 &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
 \left\{ 2C^X_{\beta  H} C^X_{\gamma L} - C^X_{\beta  L} C^X_{\gamma H}\right\} C^X_{\delta H} \;,\\
%
a_{\alpha;H}^{X;{\rm HT}} &=-\sum_\beta C_{\beta H}^X F_{\alpha\beta}^X \;,\\
%
a_{\alpha;HL}^{X;{\rm HT}} &=a_{\alpha;H}^{X;{\rm HT}} \nonumber \\
 &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
 \left\{ 2C^X_{\beta  L} C^X_{\gamma H} - C^X_{\beta  H} C^X_{\gamma L}\right\} C^X_{\delta L} \;.
%
%a_{\alpha}^{A;{\rm ET1}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^A_{\beta  H} C^A_{\gamma L} - C^A_{\beta  L} C^A_{\gamma H}\right\} C^A_{\delta H} \;,\\
%%
%a_{\alpha}^{B;{\rm ET1}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
%%
%a_{\alpha}^{B;{\rm ET2}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^B_{\beta  H} C^B_{\gamma L} - C^B_{\beta  L} C^B_{\gamma H}\right\} C^B_{\delta H} \;,\\
%%
%a_{\alpha}^{A;{\rm ET2}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;,\\
%%
%a_{\alpha}^{A;{\rm HT1}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^A_{\beta  L} C^A_{\gamma H} - C^A_{\beta  H} C^A_{\gamma L}\right\} C^A_{\delta L} \;,\\
%%
%a_{\alpha}^{B;{\rm HT1}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \;,\\
%%
%a_{\alpha}^{B;{\rm HT2}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^B_{\beta  L} C^B_{\gamma H} - C^B_{\beta  H} C^B_{\gamma L}\right\} C^B_{\delta L} \;,\\
%%
%a_{\alpha}^{A;{\rm HT2}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \;.
\end{align}
\end{subequations}
%
Here, the Fock matrix in space spanned by the intermediate and primary AO basis sets
needs to be computed. The respective contributions to ${\bf a}$ vectors can be
written in a convenient to implement form as
%
\begin{multline}
 \sum_\beta C_{\beta U}^X F_{\alpha\beta}^X = 
 \sum_\beta C_{\beta U}^X h_{\alpha\beta}^X \\
   + \sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
     \left[ P_{\delta\gamma}^{X(g)} C_{\beta U}^X - \frac{1}{2} P_{\beta\gamma}^{X(g)} C_{\delta U}^X \right] \;,
\end{multline}
%
where ${\bf h}^X$ is the one\hyp{}electron core Hamiltonian of molecule $X$.

% -----------------------
\bibliography{references}
% -----------------------

>>>>

% \mathscr{F} \approx \mathscr{F}^{A}_0 + \mathscr{F}^{B}_0 \;,
 \mathscr{F} \approx \hat{T} + 
   \sum_X^{A,B} \left\{ 
      \hat{V}^{X}_{\rm nuc} + \hat{J}({\bf P}^{X}_g) - \frac{1}{2} \hat{K}({\bf P}^{X}_g)
 \right\} \;,

@@

where $\hat{T}$ is the kinetic energy operator, 
$\hat{V}^{X}_{\rm nuc}$ is the operator associated with the electrostatic potential energy
between nuclei and electrons in molecule $X$, and the Coulomb and exchange operators
are defined by
 \hat{J}({\bf P}^{X}_g) &\equiv \sum_{\sigma\lambda} P^{X(g)}_{\lambda\sigma}
           \iint \frac{\varphi_\lambda({\bf r}_1) \varphi^*_\sigma({\bf r}_1) }{\vert {\bf r}_1 - {\bf r}_2 \vert}
           \square \; d{\bf r}_1 d{\bf r}_2 \;,\\
 \hat{K}({\bf P}^{X}_g) &\equiv \sum_{\sigma\lambda} P^{X(g)}_{\lambda\sigma}
           \iint \frac{\varphi_\lambda({\bf r}_1) \varphi^*_\sigma({\bf r}_2) }{\vert {\bf r}_1 - {\bf r}_2 \vert}
           \square \; d{\bf r}_1 d{\bf r}_2 \;.
In the above definitions, the `$\square$' symbol reminds that $\hat{J}$ and $\hat{K}$ are the integral operators
and is given by

@@

 \square \equiv \iint \Ket{{\bf r}} \Bra{{\bf r}'} \; d{\bf r} d{\bf r}' \;.

@@

Note that the kinetic energy operator cannot be decomposed into particular molecules. However, 
consider the following partitioning of the total Fock operator:
 \mathscr{F} \approx \mathscr{G}^{A}_0 + \mathscr{G}^{B}_0 \;,
in which the auxiliary operator is defined as
 \mathscr{G}^{X}_0 = \mathscr{F}^{X}_0 - \frac{1}{2}\hat{T}^{X} \;.
Here, the kinetic energy operator was symmetrically partitioned in between two molecules.
note that, without loss of generality, one can rewrite Eq.~\eqref{e:v0-et1}

@@

 \tbraket{\phi_L^A}{\left( \mathscr{G}^A_0 + \mathscr{G}^B_0 \right)}{\phi_L^B} \\

@@

 \bra{\phi_L^A} \left\{ \mathscr{G}_0^A - \hat{v}^A_{HH} \right\}

@@

 \mathscr{G}_0^B \ket{\phi_L^B} \cong \sum_{\eta\in B}^{\rm DF} V^{B;{\rm ET}}_{\eta;L} \ket{\eta} 

@@

Similar considerations apply to the hole transfer (HT) matrix elements.
The expression
%
\begin{multline}\label{e:v0-ht1}
 V^{{\rm HT1},(0)} \equiv \left< \Phi_1 \vert \mathscr{H} \vert \Phi_4 \right> \approx 
 t_{H\rightarrow L}^A \times \Big\{ 
-\big(\phi_H^A \big| \mathscr{F} \big| \phi_H^B \big) \\
   + 2 \big( \phi_H^A \phi_L^A \big| \phi_L^A \phi_H^B \big) - \big( \phi_H^A \phi_H^B \big| \phi_L^A \phi_L^A \big) 
 \Big\} 
\end{multline}
%
can be recast as
%
\begin{multline}\label{e:v0-ht1-oep}
 V^{{\rm HT1},(0)} \cong t_{H\rightarrow L}^A \Big\{ 
 \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;HL} +
 \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;H}
 \Big\} \;,
\end{multline}
%
with the effective potential matrices given by
%
%
\begin{subequations}
\begin{align}
 \bra{\phi_H^A} \left\{ -\mathscr{G}_0^A - \hat{v}^A_{LL} \right\}
 + \bra{\phi_L^A} 2 \hat{v}^A_{HL} 
 &\cong \sum_{\xi\in A}^{\rm DF} \bra{\xi} V^{A,{\rm HT}}_{\xi,HL} \\
 -\mathscr{G}_0^B \ket{\phi_H^B} &\cong \sum_{\eta\in B}^{\rm DF} V^{B,{\rm HT}}_{\eta,H} \ket{\eta} 
\end{align}
\end{subequations}
%
The twin term HT2 is accordingly
%
\begin{multline}\label{e:v0-ht2-oep}
 V^{{\rm HT2},(0)} \cong t_{H\rightarrow L}^B \Big\{ 
 \sum_{\eta\in B}^{\rm DF} S_{\eta H}^{BA} V^{B;{\rm HT}}_{\eta;HL} +
 \sum_{\xi\in A}^{\rm DF} S_{\xi H}^{AB} V^{A;{\rm HT}}_{\xi;H}
 \Big\} \;.
\end{multline}
%
As can be seen, to compute Hamiltonian matrix elements
associated with the electron and hole transfer, eight in total
different OEP matrices need to be pre\hyp{}computed and stored in a file.
Note that each of these matrices is actually just a vector of length
equal to the size of the auxiliary basis set chosen. Therefore,
it can be predicted that the computational cost of evaluation of the OEP\hyp{}based Hamiltonian
matrix elements is negligible as compared to the original TI/CIS method.
In Appendix~\ref{a:gdf-formulae}, explicit working formulae
for the OEP matrix elements are given in terms of the AO's.

@@

As developed in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.XXXX}, 
the matrix elements of the OEP operators
are computed from the generalized density fitting scheme, according to which
\begin{equation}\label{e:a:voep}
 V^X_{\xi, M} = \sum_{\xi'\in X}^{\rm DF} 
                \sum_{\varepsilon\in X}^{\rm RI}
                [{\bf R}^{-1}]_{\xi\xi'} R_{\xi'\varepsilon} H^{X;M}_{\varepsilon; N} \;,
 H^{X;M}_{\varepsilon; N} = \sum_{\varepsilon'\in X}^{\rm RI}
                        [{\bf S}^{-1}]_{\varepsilon\varepsilon'} a_{\varepsilon';N}^{X; M} \;.
In the above equations, the auxiliary matrices are defined by
 S_{\eta\xi}  = \int 
                       \varphi^*_\eta({\bf r}) \varphi_\xi({\bf r}) 
                 d{\bf r} 
 R_{\eta\xi}  = \iint 
                       \frac{ \varphi^*_\eta({\bf r}_1) \varphi_\xi({\bf r}_2) } 
                            {\vert {\bf r}_1 - {\bf r}_2\vert}  
The elements of vector ${\bf a}^{X;M}_N$ for $M=$ ET1, ET2, HT1 and HT2 and $N=$ H, L and HL are given by
\begin{subequations}\label{e:a:aoep}
a_{\alpha;L}^{X;{\rm ET}} &= \sum_\beta C_{\beta L}^X G_{\alpha\beta}^X \;,\\
a_{\alpha;HL}^{X;{\rm ET}} &= a_{\alpha;L}^{X;{\rm ET}} \nonumber \\
 &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
 \left\{ 2C^X_{\beta  H} C^X_{\gamma L} - C^X_{\beta  L} C^X_{\gamma H}\right\} C^X_{\delta H} \;,\\
a_{\alpha;H}^{X;{\rm HT}} &=-\sum_\beta C_{\beta H}^X G_{\alpha\beta}^X \;,\\
a_{\alpha;HL}^{X;{\rm HT}} &=a_{\alpha;H}^{X;{\rm HT}} \nonumber \\
 &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
 \left\{ 2C^X_{\beta  L} C^X_{\gamma H} - C^X_{\beta  H} C^X_{\gamma L}\right\} C^X_{\delta L} \;.
%a_{\alpha}^{A;{\rm ET1}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^A_{\beta  H} C^A_{\gamma L} - C^A_{\beta  L} C^A_{\gamma H}\right\} C^A_{\delta H} \;,\\
%%
%a_{\alpha}^{B;{\rm ET1}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \;,\\
%%
%a_{\alpha}^{B;{\rm ET2}} &= \sum_\beta C_{\beta L}^B F_{\alpha\beta}^B \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^B_{\beta  H} C^B_{\gamma L} - C^B_{\beta  L} C^B_{\gamma H}\right\} C^B_{\delta H} \;,\\
%%
%a_{\alpha}^{A;{\rm ET2}} &= \sum_\beta C_{\beta L}^A F_{\alpha\beta}^A \;,\\
%%
%a_{\alpha}^{A;{\rm HT1}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^A_{\beta  L} C^A_{\gamma H} - C^A_{\beta  H} C^A_{\gamma L}\right\} C^A_{\delta L} \;,\\
%%
%a_{\alpha}^{B;{\rm HT1}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \;,\\
%%
%a_{\alpha}^{B;{\rm HT2}} &=-\sum_\beta C_{\beta H}^B F_{\alpha\beta}^B \nonumber \\
% &+\sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
% \left\{ 2C^B_{\beta  L} C^B_{\gamma H} - C^B_{\beta  H} C^B_{\gamma L}\right\} C^B_{\delta L} \;,\\
%%
%a_{\alpha}^{A;{\rm HT2}} &=-\sum_\beta C_{\beta H}^A F_{\alpha\beta}^A \;.
Here, the auxiliary one\hyp{}electron ${\mathscr{G}}^X$ operator represented
in space spanned by the intermediate and primary AO basis sets
needs to be computed. The respective contributions to ${\bf a}$ vectors can be
written in a convenient to implement form as
 \sum_\beta C_{\beta U}^X G_{\alpha\beta}^X = 
 \sum_\beta C_{\beta U}^X \left( \frac{1}{2} T_{\alpha\beta}  + V_{{\rm nuc};\alpha\beta}^X \right)\\
   + \sum_{\beta\gamma\delta} \braket{\alpha\beta}{\gamma\delta}
     \left[ P_{\delta\gamma}^{X(g)} C_{\beta U}^X - \frac{1}{2} P_{\beta\gamma}^{X(g)} C_{\delta U}^X \right] \;,
 T_{\alpha\beta} &= -\frac{1}{2} \int \varphi^*_\alpha({\bf r}) \nabla^2_{{\bf r}} \varphi_\beta({\bf r})\; d{\bf r} 
 \text{ and}\\ 
 V_{{\rm nuc};\alpha\beta}^X &= -\sum_{x\in X} Z_x \int 
           \frac{\varphi^*_\alpha({\bf r})  \varphi_\beta({\bf r}) }{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;.
In the above, $Z_x$ is the atomic number of the $x$th atom and $\nabla^2_{{\bf r}}\equiv 
\frac{\partial^2}{\partial x^2} + \frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2}$.
% -----------------------
\bibliography{references}
% -----------------------
